---
output:
  html_document:
    keep_md: yes
---

# Practical Machine Learning
## Course Project
#### January 2015

The goal of this project is to use the data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants to quantify how well the participant performs barbell lifts. The participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The data for this analysis came from "Qualitative Activity Recognition of Weight Lifting Exercises" by Velloso et al. The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


### Download csv training and testing data
```{r, echo=TRUE}
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
```

### Read in the CSV data
```{r, echo=TRUE}
pml_train <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
pml_test <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
dim(pml_train)
dim(pml_test)
```

### Remove any columns that contain only missing values. 100 columns are removed from both sets.
```{r, echo=TRUE}
pml_train <- pml_train[,colSums(is.na(pml_train)) == 0]
pml_test <- pml_test[,colSums(is.na(pml_test)) == 0]
dim(pml_train)
dim(pml_test)
```

### Keep only the columns that contain the words belt, arm, dumbbell, and forearm in the name. This becomes all columns except for the first 7 columns.
```{r, echo=TRUE}
names(pml_train)
pml_train <- pml_train[,-c(1:7)]
pml_test <- pml_test[,-c(1:7)]
dim(pml_train)
dim(pml_test)
```

### Partition the training data into a training set (80%) and a validation set (20%)
```{r, echo=TRUE}
library(caret)
set.seed(123)
trainingset <- createDataPartition(pml_train$classe, p = 0.8, list = FALSE)
training <- pml_train[trainingset, ]
validation <- pml_train[-trainingset, ]
dim(training)
dim(validation)
```


## Random Forest Model
```{r, echo=TRUE}
library(randomForest)
rf <- randomForest(classe ~ ., data = training, importance = TRUE, ntrees = 5)
```

### Check the accuracy of the training set when using the Random Forest model
```{r, echo=TRUE}
predict_training <- predict(rf, training)
print(confusionMatrix(predict_training, training$classe))
```

### Now, check the accuracy of the Random Forest model against the validation set
```{r, echo=TRUE}
predict_validation <- predict(rf, validation)
print(confusionMatrix(predict_validation, validation$classe))
```



### Let's use the model on the testing data set to determine the outcome of classe
```{r, echo=TRUE}
predict_test <- predict(rf, pml_test)
predict_test
```
